Article Title

### 2013-10-31 sections/Figure3.tex

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
\caption{
In simulation, selection amongst estimators recognizes the true covariance atructure. 
{\bf A--D} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations. Matrix A has no low-dimensional structure. Matrices B--D have low-dimensional structture corresponding to the respective graphical models in Fig.~2B--D.
{\bf E--H} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
{\bf I--L:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.  Excess loss $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ is only accessible with the true covariance matrix $\Sigma$ is known and zero excess loss implies absolutely correct estimation. Estimators that can represent the graphical model outperform those that cannot.  Error bars indicate the standard deviation. 
{\bf M--P:} Performance evaluation using cross-validated \emph{empirical loss} from Eq.~\ref{eq:empiricalLoss} without knowledge of true $\Sigma$. Cross-validation reproduces the same relationship between performances of the estimators.
}\label{fig:03}
\end{figure}

>>>>

\begin{figure}[htp]
\centering
\includegraphics[width=1.0\textwidth]{figures/Figure3.pdf}
\caption{
Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper. It assumed that all interactions are linear, \emph{i.e.}\;the mean firing rate of any neuron can be computed by a linear combination of the firing rates of other neurons and latent units.
\textsf{A}: ``Independent.'' A population with no interactions corresponds to the diagonal covariance matrix.
\textsf{B}: ``Latent factors.'' Observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
\textsf{C}: ``Sparse.'' Partial correlations between a subset of pairs of observed neurons, zero partial correlations between all other pairs.  
\textsf{D}: ``Sparse+latent.''  Partial correlations between a subset of pairs of observed neurons and a few latent factors interacting with the entire population. 
{\sf E--H} are examples of true $100\times100$ covariance matrices of multivariate normal distributions used in simulations. Matrix A has no low-dimensional structure. Matrices B--D have low-dimensional structture corresponding to the respective graphical models in Fig.~2B--D.
{\sf I--L} are examples of sample covariance matrices obtained from samples of size $n=1000$ from distributions with respective true covariance matrices A--D.
{\sf M--P:} Performance evaluation of covariance estimators A, B, C, D from samples obtained from models A, B, C, D.  Excess loss $\loss{\hat\Sigma,\Sigma}-\loss{\Sigma,\Sigma}$ is only accessible with the true covariance matrix $\Sigma$ is known and zero excess loss implies absolutely correct estimation. Estimators that can represent the graphical model outperform those that cannot.  Error bars indicate the standard deviation. 
{\sf Q--T:} Performance evaluation using \emph{validated loss} from Eq.~\ref{eq:validationLoss} without knowledge of true $\Sigma$. Cross-validation reproduces the same relationship between performances of the estimators.
}\label{fig:03}
\end{figure}

### 2013-10-31 sections/Figure4.tex

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
\caption{{\bf Estimator $\mathcal D$ (sparse+full-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
{\bf A:}  Histogram of improvements in validation loss by covariance estimator $\mathcal D$ relative to the samle covariance matrix from the 31 imaged sites used in the study.  The vertical red line denotes the median improvment. 
{\bf B:} Improvements relative to estimator $\mathcal A$ (shrinkage toward the diagonal).
{\bf C:} Improvements relative to estimator $\mathcal B$ (shrinkage toward the factor model).
{\bf D:} Improvements relative to estimator $\mathcal C$ (sparse inverse covariance).
}\label{fig:04}
\end{figure}

>>>>

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure4.pdf}
\caption{{\bf Estimator $\mathcal D$ (sparse+low-rank inverse covariance) dominates in dense recordings from primary visual cortex.}
{\bf A:}  Histogram of improvements in validation loss by covariance estimator $\mathcal D$ relative to the samle covariance matrix from the 31 imaged sites used in the study.  The vertical red line denotes the median improvment. 
{\bf B:} Improvements relative to estimator $\mathcal A$ (shrinkage toward the diagonal).
{\bf C:} Improvements relative to estimator $\mathcal B$ (shrinkage toward the factor model).
{\bf D:} Improvements relative to estimator $\mathcal C$ (sparse inverse covariance).
}\label{fig:04}
\end{figure}

### 2013-10-31 sections/Results.tex
minor

In the second regularized estimator, the target estimate is the factor model $\hat F = \hat L \hat L^\T + \hat \Psi$ with $d$ factors so that $\hat L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\hat \Psi$ contains the independent variances of each neuron.
Then the estimate is 
\begin{equation}
\hat \Sigma_{\sf factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
\end{equation}
This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $\hat F$ expresses the assumption that the fluctuations in population activity are driven by relatively small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:03}B).

@@

Finally, we consider a fourth estimator, which allows for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:03}D). This estimator has the form
\begin{equation}
\hat\Sigma_{\sf sparse+latent} = (\hat S - \hat L\hat L^\T)^{-1},
\end{equation}
where, as above, $\hat S$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of the sparsity of local interactions $\rho$.

\subsection*{Simulation}

To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different underlying structures. Each population contained 100~neurons. The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), low-rank (Fig.~\ref{fig:03}B), sparse inverse (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (~\ref{fig:03}D). \Acomment{Integrate schematics from Fig.~1 into this figure (make it the first row) for better correspondence} In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).

@@

Thus, if taken in isolation a regularized estimator improves the estimate we should not interpret this result to suggest that the estimator's target has the same low-dimensional structure as the data-generating process. However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error. Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~2\,A--D, fourth row). Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~2\,B,\,C). \Acomment{May want to include a smaller sample size as well since here the more reduced estimators (B, C) may actually outperform estimator D when the population structure matches their target and there is little data (say, 1.5x number of neurons)}

Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on toy data. However, an unbiased empirical estimate of the excess loss exists (see Methods, Eq.~\ref{eq:validationLoss} \Acomment{This may need a brief sentence about the intuition what's the difference, or a reference to the relevant methods section}).  To verify that our approach should also work in a realistic situation without access to ground truth, we also computed the empirical loss. Indeed, without access to ground truth, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~2, last row). Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, empirical loss is a much noisier measurement than excess loss. In addition, it does not converge to zero with increasing sample size as excess loss does.

\subsection*{Covariance estimation on neural data}

We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}. We presented numerous repetitions of full-field drifting gratings (Fig.~\ref{fig:01}A and \ref{fig:01}B) to the eye contralateral to the imaged site. This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4. We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove effects of the stimulus. From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).

@@

As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~4) \Acomment{I think we should show that}. Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator? I've seen many people do this and it provides an ordering. I feel like we should come up with a way of ordering them somehow, even if it's under certain assumptions that may not be entirely correct. I'm pretty sure if we don't do it the reviewers will bring it up anyway...}. Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data. The improvement from the low-rank estimator to the combined one is much larger than that from the sparse inverse to the combined one, suggesting that in this dataset direct interactions contribute more strongly to the correlation structure than hidden units do.

\subsection*{Relationship between functional covariance structure and circuit architecture}

>>>>

In the second regularized estimator, the target estimate is the factor model $\hat F = \hat L \hat L^\T + \hat \Psi$ with $d$ factors so that $\hat L$ is the $p\times d$ matrix of \emph{factor loadings} and the diagonal matrix $\hat \Psi$ contains the independent variances of each neuron.
Then the estimate is 
\begin{equation}
\hat \Sigma_{\sf factor} = (1-\lambda) \hat \Sigma_0 + \lambda \hat F,
\end{equation}
This estiamtor has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $\hat F$ expresses the assumption that correlated fluctuations in population activity are driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (Fig.~\ref{fig:03}B).

@@

Finally, we consider a fourth estimator, which allows for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:03}D). This estimator has the form
$\rho$.
\begin{equation}
\hat\Sigma_{\sf sparse+latent} = (\hat S - \hat L\hat L^\T)^{-1},
\end{equation}
where, as above, $\hat S$ is a sparse matrix and $\hat L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of local interactions 
\subsection*{Simulation}

To verify our approach and to illustrate the performance of the four regularized estimators, we constructed five model populations with different covariance structures. Each population contained 50~neurons. 
The first four populations matched the low-dimensional structure of our four estimators: independent (Fig.~\ref{fig:03}A), latent factors (Fig.~\ref{fig:03}B), sparse partial correlations (Fig.~\ref{fig:03}C) and low-rank combined with sparse inverse (Fig.~\ref{fig:03}D).  
In addition, we considered a fifth population that had no low-dimensional structure (Fig.~\ref{fig:03}\,E).

@@

Thus, if taken in isolation a regularized estimator improves the estimate we should not interpret this result to suggest that the estimator's target has the same low-dimensional structure as the data-generating process. However, when comparing multiple estimators against each other, the one whose target estimate most closely matches the true value with the smallest number of parameters will reduce the estimation error with the least increase in approximation error. Indeed, in all four toy examples with a low-dimensional structure, the estimator with the matching regularization target performed best (Fig.~\ref{fig:03}\,A--D, fourth row). 
Since the estimator combining low-rank and sparse interactions combines to low-dimensional targets and includes the simpler estimators as special cases, it performed almost as well even when the low-dimensional structure did not include either a low-rank component or sparse interactions (Fig.~\ref{fig:03}\,B,\,C).

Since the above evaluations of excess loss require knowledge of ground truth, this analysis can be done only on simulated data. However, \emph{validation loss} can be used as an unbiased empirical estimate of loss  (see Methods, Eq.~\ref{eq:validationLoss}. Validation loss uses an independent validation sample to estimate the value of loss without access to ground truth.   Indeed, this analysis revealed a pattern of results that reproduced the results obtained with access to ground truth above (Fig.~\ref{fig:03}, last row). Because validation loss is computed by comparing estimates to noisy sample covariance matrices from smaller validation sets, validation loss has higher variance than excess loss. Under our chosen loss function, validation loss does not converge to zero.

\subsection*{Covariance estimation in neural data}

We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Stosiek:2003,Reddy:2005}. We presented numerous repetitions of full-field drifting gratings (Fig.~\ref{fig:01}A and \ref{fig:01}B) to the eye contralateral to the imaged site. This technique allowed us to record from a large number (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4. We deconvolved somatic calcium signals using sparse nonnegative deconvolution \cite{Vogelstein:2010} (Fig.~\ref{fig:01}C and \ref{fig:01}D) and subtracted the average stimulus response to remove averaged linear effects of the stimulus. From this residual response we computed the sample noise covariance matrix (Fig.~\ref{fig:01}E).

@@

As expected, all regularized estimators outperformed the sample covariance estimator substantially (Fig.~\ref{fig:04}) \Acomment{I think we should show that}. Say something about how shrinkage, low-rank and sparse inverse relate \Acomment{Why do we do pairwise comparisons instead of just showing median (across sites) log loss relative to the combined estimator? I've seen many people do this and it provides an ordering. I feel like we should come up with a way of ordering them somehow, even if it's under certain assumptions that may not be entirely correct. I'm pretty sure if we don't do it the reviewers will bring it up anyway...}. Finally, the combined sparse and low-rank estimator dominated all others significantly (Fig.~4), showing that both hidden units and direct interactions are important in our data. The improvement from the low-rank estimator to the combined one is much larger than that from the sparse inverse to the combined one, suggesting that in this dataset direct interactions contribute more strongly to the correlation structure than hidden units do.

\input{sections/Figure4.tex}

\subsection*{Relationship between functional covariance structure and circuit architecture}
\input{sections/Figure5.tex}
\input{sections/Figure6.tex}
